---
title: "Lab4"
author: "Your Name"
date: "2024-09-26"
output: "github_document"
---

```{r}

library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)

```

# Load dataset
```{r}

lab4data <- read.csv("/Users/jostarenky/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot
ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + facet_wrap(~Group)+theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + facet_wrap(~Group)+theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + facet_wrap(~Group)+theme_classic()

qq+ geom_qq_line()

#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)

#shapiro test: not normal bc p is significant
#skew and kurtosis: normal


```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group
describeBy(Performance ~ Group, data = lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic, p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?
This variable is not normal because even though the skew and kurtosis values fit the normal ranges, the Shapiro test and visualizations indicate that the variable is not normal


# Equal Variance between Groups
## Descrptive Variance
```{r}
#get rid of missing values
lab4data_clean<-drop_na(lab4data)

var(lab4data_clean$Performance)


lab4data_clean %>%
  group_by(Group) %>%
  summarize(variance = var(Performance))


```

## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)
#cannot run MANOVA bc can't install bruce package
#MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance~Group, lab4dataConG1)


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?
No, because the result was significant
# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)



```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 
Shapiro test is now not significant, indicating normality. Equal variance between groups is now true. Levene's test is not significant, so equal variance across groups

```{r}
#visualizations
ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance_log)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()



#Normality tests
describe(lab4data$Performance_log)

shapiro.test(lab4data$Performance_log)

#Use the describeBy() function to get skewness and kurtosis by group
describeBy(Performance_log ~ Group, data = lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance_log)$statistic, p_value = shapiro.test(Performance_log)$p.value)

#Descriptive variance
lab4data_clean<-drop_na(lab4data)

var(lab4data_clean$Performance_log)


lab4data_clean %>%
  group_by(Group) %>%
  summarize(variance = var(Performance_log))

#Equal variance across groups
leveneTest(Performance_log~Group, lab4data)
#cannot run MANOVA bc can't install bruce package
#MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance_log~Group, lab4dataConG1)


```
